---
title: "Figures for Paper"
author: "Anna-Leigh Brown"
date: "06/10/2020"
output: 
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
---


```{css, echo = FALSE}
h1, #TOC>ul>li {
  color: #000000;
    font-weight: bold;
}

h2, #TOC>ul>ul>li {
  color: #41494D;
}

```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
knitr::opts_chunk$set(
  comment = '', fig.width = 9, fig.height = 9
)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here",
       "ggthemes",
       "ggrepel")
```


# iPSC splicing

Read in the splicing results returned by MAJIQ and make a volcano plot, only highlight 
genes of interest with a label.

```{r volcano_plots}

ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
ipsc_de = fread(file.path(here::here(),"data","ipsc_differential_expression.csv"))


splicing_dots_tables <- ipsc_splicing %>% 
  mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ gene_name,
                                    T ~ "")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ junction_name,
                                    T ~ "")) 

fig1a = ggplot() +
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions != 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 10) + 
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 10) + 
  geom_text_repel(data = splicing_dots_tables[junction_name != ""],
                  aes(x = deltaPSI, y =log10_test_stat,
                      label = label_junction,
                      color = as.character(de_novo_junctions)), point.padding = 0.3,
                    nudge_y = 0.2,
                  min.segment.length = 0,
                  box.padding  = 2,
                  size=6,show.legend = F) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  geom_vline(xintercept = -0,linetype="dotted") + 
     scale_fill_manual(name = "",
                      breaks = c("Annotated Junction", "Novel Junction"),
                      values = c("Annotated Junction" = "#648FFF", "Novel Junction" = "#fe6101") ) +
       scale_color_manual(name = "",
                      breaks = c("0", "1"),
                      values = c("1" = "#648FFF", "0" = "#fe6101") ) +
  guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab(expression(paste("-Lo", g[10], " Test Statistic"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
  xlim(-1,1)


de_table = ipsc_de %>% 
  mutate(contains_cryptic = gene_name %in% splicing_dots_tables[cryptic_junction == T,unique(gene_name)]) %>% 
  mutate(contains_cryptic = as.character(as.numeric(contains_cryptic))) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2","TARDBP") ~ gene_name,
                                    T ~ "")) %>% 
  mutate(graph_alpha = ifelse(padj < 0.1, 1, 0.2)) %>% 
  mutate(y_data = -log10(padj))
  
  
fig1b = ggplot(data = de_table) + 
  geom_point(data = de_table %>% filter(contains_cryptic == "0"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 10) + 
    geom_point(data = de_table %>% filter(contains_cryptic == "1"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 10)+ 
    geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
                  aes(x = log2FoldChange, 
                      y = -log10(padj),
                      label = label_junction,
                      color = as.character(contains_cryptic)),
                   nudge_y = 5,
                  min.segment.length = 0,
                  box.padding  = 4,
                  size=6,show.legend = F) +
       scale_fill_manual(name = "",
                      breaks = c("No Cryptic", "Contains Cryptic"),
                      values = c("No Cryptic" = "#648FFF", "Contains Cryptic" = "#fe6101") ) +
         scale_color_manual(name = "",
                      breaks = c("1", "0"),
                      values = c("1" = "#fe6101", "0" = "#648FFF") ) +
  xlim(-7.5,7.5) + 
  ylab(expression(paste("-Lo", g[10], " P-value"))) +
    guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab(expression(paste("Lo", g[2], " Fold Change"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
    geom_hline(yintercept = -log10(0.1))  +
  geom_vline(xintercept=c(0), linetype="dotted")

print(fig1a)
print(fig1b)
```



# UNC13A CE PSI across TDP-43 KD experiments

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
The SK-N-DZ cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
There was variability on the Klim hMN set on allelic expression. 

```{r read in }

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
```


## barplot UNC13A CE PSI - full five

```{r}
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    scale_y_continuous(labels = scales::percent) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
```

## barplot UNC13B NMD PSI

```{r}
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "unc13b_nmd_exon_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B \nNMD Exon PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
```

## TARDBP RNA and UNC13A Cryptic

```{r}



rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```

## UNC13A RNA and UNC13A Cryptic

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12,
             method = 'spearman',
           cor.coef.name = 'rho') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
      scale_y_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
```
## UNC13A RNA and UNC13A IR
```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```
## TARDBP RNA and UNC13A IR


```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```

## barplot UNC13A relative RNA across experiments

```{r}
rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "cryptic_psi_full",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("UNC13A CE PSI") + 
  xlab("") + 
  guides(color = FALSE) 
```

```{r}
stmn2 = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "stmn_2_cryptic_psi",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("STMN2 Cryptic PSI") + 
  xlab("") + 
  guides(color = FALSE)

####barplot UNC13B normalized IR####
unc13b_ir = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13b_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13B \nIntron Retention Ratio") + 
    xlab("") + 
    guides(color = FALSE) + 
      theme(legend.position = "none") +
  theme(text = element_text(size = 18)) +
      guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
####barplot UNC13A normalized IR####
unc13a_ir = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13A \nIntron Retention Ratio") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

####scatterplot stmn2 normalized TARDBP####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = stmn_2_cryptic_psi, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
    theme(legend.position = 'bottom') 
####scatterplot UNC13BIR normalized TARDBP####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
####scatterplot UNC13AIR and UNC13A Crptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13a_ir)) +
    theme(legend.position = 'bottom')
####scatterplot UNC13BIR and UNC13A Crptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)
```

## scatterplot UNC13BIR and UNC13B NMD

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)

```

# NMD experiment on sy5y cells

```{r}

fbp_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
fbp_sy$condition <- as.factor(fbp_sy$condition)
fbp_sy$condition <- factor(fbp_sy$condition,levels(fbp_sy$condition)[c(2,1)])

stattest_sy <- fbp_sy %>%
    group_by(gene) %>%
    t_test(value ~ 1, mu = 1) %>%
    add_significance() %>%
    add_xy_position(x="gene")
stattest_sy

nmd_plot_sy <- ggbarplot(fbp_sy,
                     x = 'gene',
                     y = 'value',
                     add = c("mean_se","jitter"),
                     fill = 'condition',
                     color = 'condition',
                     position = position_dodge(0.8),
                     dot.size = 10) +
    scale_y_continuous(labels = scales::percent) + 
    ggpubr::theme_pubr() + 
    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) + 
    scale_color_manual(values = c("#1C2617","#262114")) +
    ylab("Percent of transcript after CHX treatment") +
    xlab("") +
    stat_pvalue_manual(stattest_sy, label="p.signif", y.position = 9)
nmd_plot_sy

```

```{r}

fbp_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
fbp_dz$condition <- as.factor(fbp_dz$condition)
fbp_dz$condition <- factor(fbp_dz$condition,levels(fbp_dz$condition)[c(2,1)])

stattest_dz <- fbp_dz %>%
    group_by(gene) %>%
    t_test(value ~ 1, mu = 1) %>%
    add_significance() %>%
    add_xy_position(x="gene")
stattest_dz

nmd_plot_dz <- ggbarplot(fbp_dz,
                        x = 'gene',
                        y = 'value',
                        add = c("mean_se","jitter"),
                        fill = 'condition',
                        color = 'condition',
                        position = position_dodge(0.8),
                        dot.size = 10) +
    scale_y_continuous(labels = scales::percent) + 
    ggpubr::theme_pubr() + 
    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) + 
    scale_color_manual(values = c("#1C2617","#262114")) +
    ylab("Percent of transcript after CHX treatment") +
    xlab("") +
    stat_pvalue_manual(stattest_dz, label="p.signif", y.position = 26)
nmd_plot_dz

```

# iPSC protein abundance mass spec

```{r}
peptides = fread(file.path(here::here(),"data","peptide_abdundance.csv"))

peptides %>% 
  mutate(gene = fct_relevel(gene, "UNC13A","UNC13B")) %>% 
  ggbarplot(,
            x = "condition",
            add = c("mean_se"),
            y = "protein",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8),
            facet.by = 'gene') + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("Protein abundance") + 
  xlab("") + 
  geom_jitter(pch = 21,height = 0,aes(fill = condition),size = 2) + 
  guides(color = FALSE) +
  facet_wrap(~gene, scales = 'free') +
  theme(legend.position = 'none') + 
  scale_y_continuous(labels = scientific) + 
  stat_compare_means(comparisons = list(c("Control","TDP-43 KD")),
                     label = 'p.signif',tip.length = 0,
                     size = 10)

 
```


# Patient summary statistics


```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(rs12973192) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) 

print(glue::glue("Number of unique patients: {clean_data_table[,length(unique(participant_id))]}"))
print(glue::glue("Number of unique tissue samples: {clean_data_table[,length(unique(sample))]}"))
print("Patients Per Disease Category")
clean_data_table[,length(unique(participant_id)),by = disease]
print("Tissues Per Disease Category")
clean_data_table[,length(unique(sample)),by = disease]
print("Number of patients per UNC13A SNP genotype")
unique(clean_data_table[,.(participant_id,rs12608932,rs12973192)]) %>% select(-participant_id) %>% table()
print("Number of tissues per disease")
clean_data_table[,.N,by = c("disease","tissue_clean")]
print("Number of partcipants by mutation and  disease")
clean_data_table[,length(unique(participant_id)),by = c("mutations","disease")]

print(glue::glue("Number of patients per pathology:"))
clean_data_table[,length(unique(participant_id)),by = .(pathology)]
```
# UNC13A cryptic is an event that is specific to TDP-43 proteinopathy

FTLD-non-TDP are those with TAU and FUS aggregates

Non-tdp ALS are those with SOD1 or FUS mutations. The n's are quite low on this 
unfortunately, only 8 ALS with SOD1 and 2 with FUS mutations.  

First we look at detection rate in tissues affected by TDP-43 proteinopathy,
For FTLD this is frontal and temporal Cortices, and
for ALS this is lumbar, cervical, and thoracic spinal cord samples as well as 
motor cortex. For controls we also take all 6 tissues, 
frontal,temporal,motor cortices and the lumbar, cervical, and thoracic spinal cords. 

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)

```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_1",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
## Pecent of patients UNC13A detected
Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.

```{r}

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 24)) + 
  xlab("N individuals")
```

## Pecent of samples UNC13A detected by tissue


```{r}
boxplot_table %>% 
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(sample,disease_group2,detected,tissue_clean) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  ungroup() %>% 
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>% 
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = detection_name),position = position_dodge2(width = 0.8, preserve = "single")) + 
  ggpubr::theme_pubr() + 
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N tissues") + 
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) + 
  theme(axis.text.x =  element_text(size = 14)) + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  +
  scale_fill_manual(values = c("#4C97B5","#4C97B5","#4C97B5","#4C97B5","#4C97B5","red","#C87156","#C87156")) +
  theme(legend.position = "none") 
```

## Pecent of samples UNC13A detected by tissue and genotype

```{r}

boxplot_table %>% 
  left_join(clean_data_table %>% dplyr::select(sample, rs12973192)) %>% 
  unique() %>% 
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(sample,disease_group2,detected,tissue_clean,rs12973192) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean,rs12973192) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  ungroup() %>% 
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>% 
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = rs12973192),position = position_dodge2(width = 0.8, preserve = "single")) + 
  ggpubr::theme_pubr() + 
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N tissues") + 
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) + 
  theme(axis.text.x =  element_text(size = 14)) + 
      theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  +
  theme(legend.position = "bottom") 
```

## UNC13A CE level across tissue

This only plots the short novel acceptor and novel donor

```{r onlyshort}
boxplot_table %>% 
    filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum") )%>% 
    group_by(disease_group2,tissue_clean) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    ungroup() %>% 
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, 
               y = inclusion_reads * 10^6,
               fill ="red")) + 
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1)) + 
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom") 
```


## UNC13A CE level across tissue - split by junction type

```{r}
clean_data_table %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_1|_5prime_2",variable)) %>% 
      # filter(grepl("_5prime_2",variable)) %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  dplyr::select(sample,disease_group2,
                tissue_clean,
                junction_name,
                value) %>% 
    unique() %>% 
    filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum"))%>% 
    group_by(disease_group2,tissue_clean) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    ungroup() %>% 
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, 
               y = value * 10^6,
               fill = junction_name)) + 
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single")) + 
    geom_point(size = 2,pch = 21, alpha = 0.7, position = position_jitterdodge(jitter.width = 0.2))+ 
    scale_y_log10() + 
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
        theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  +
  theme(legend.position = "bottom") 
```


## UNC13A CE level across disease types


```{r}


disease_comparisons = list( c("Control","ALS-TDP"),
                           c("Control","ALS \n non-TDP"),
                           c("Control","FTLD-TDP"),
                           c("Control","FTLD \n non-TDP" ))

table_to_test = boxplot_table %>% 
    filter(disease_tissue == T) %>% 
    group_by(disease_group2) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
    dplyr::select(detection_name,
                  participant_id,
                  inclusion_reads,
                  disease_group2,
                  tissue_clean,
                  disease_tissue) %>% 
    unique() 

test_pair = pairwise.wilcox.test(table_to_test$inclusion_reads, table_to_test$detection_name,
                     p.adjust.method = "BH") %>% broom::tidy()

test_pair = test_pair %>% 
  mutate(p_value_draw = case_when(p.value < 0.0001~ "***",
                                  p.value < 0.01 ~ "**",
                                   p.value < 0.05 ~ "*",
                          TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>% 
  mutate(y.position = seq(0.25,by = 0.1,length.out = 7))

table_to_test %>% 
    ggplot(aes(x = detection_name, y = inclusion_reads * 10^6)) + 
    geom_boxplot() + 
    geom_jitter(height = 0) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    ylab("UNC13A cryptic inclusion \n reads per million") + 
    theme(text = element_text(size = 24)) + 
    xlab("N samples") + 
    stat_pvalue_manual(test_pair %>% filter(p.value < 0.05),
                       label = "p_value_draw",size = 8) +
   stat_compare_means(size = 8)

```


# UNC13A RNA expression (TPM) 

## UNC13A TPM by disease and tissue


```{r}
comps = unique(clean_data_table$disease_group2) %>% gsub("Control", " Control",.) %>% 
  combn(.,2,simplify = F)

clean_data_table %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
 filter(tissue_clean %in% 
           c("Frontal Cortex", 
             "Lumbar Spinal Cord", 
             "Cervical Spinal Cord",
             "Motor Cortex", 
             "Temporal Cortex",
             "Cerebellum") )%>% 
mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                              "Frontal Cortex",
                              "Lumbar Spinal Cord",
                              "Temporal Cortex",
                              "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) + 
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none')
```

## Correlation STMN2 Cryptic PSI and UNC13A TPM 

only in ALS/FTLD-TDP

```{r}
clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  ylim(0,50)
```

## Correlation STMN2 Cryptic PSI and UNC13A TPM - tissue separations


```{r}

clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = "free_y")
```

## Correlation between UNC13A CE PSI and UNC13A TPM 


```{r}
clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13B_TPM)) + 
  geom_point() + 
  stat_cor(label.y = 38,size = 8,method = 's',cor.coef.name = 'rho') + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 22)) +
  scale_x_continuous(labels = scales::percent) 

```

## Correlation between UNC13A CE PSI and UNC13A TPM - tissue separated

```{r}
clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = 'free') +
  scale_x_continuous(labels = scales::percent)  

```




# Relationship between STMN2 and UNC13A CE PSI 
## Scatter plot showing the correlation in non-log space in cortex UNC13A & STMN2 CE

```{r}
clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) + 
  geom_point() + 
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent)  +
  scale_x_continuous(labels = scales::percent) 
```

## Side by side comparison UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>% 
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  # filter(unc13a_cryptic_leaf_psi > 0) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  melt() %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>% 
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable)) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  scale_color_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean,nrow = 3,scales = 'free') +
    scale_y_continuous(labels = scales::percent)  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')

```

## Side by side comparison UNC13A & STMN2 CE PSI v2 
```{r}
clean_data_table %>% 
  filter(UNC13A_annotated_leaf > 10) %>% 
  filter(STMN2_annotated_leaf > 10) %>% 
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  # filter(unc13a_cryptic_leaf_psi > 0) %>%
  # filter(disease_tissue == T) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  filter(tissue_clean %in% c("Motor Cortex", "Cervical Spinal Cord", "Frontal Cortex", "Lumbar Spinal Cord", "Temporal Cortex")) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_rat = log2((unc13a_cryptic_leaf_psi + 0.1) / (stmn_2_cryptic_psi_leaf + 0.1))) %>% 
  group_by(tissue_clean, disease_group2) %>% 
  mutate(average_log2Fold = mean(log2_rat)) %>% 
  ungroup() %>% 
  dplyr::select(-log2_rat) %>% 
  pivot_longer(cols = c("stmn_2_cryptic_psi_leaf","unc13a_cryptic_leaf_psi")) %>% 
  mutate(tissue_clean = fct_reorder(tissue_clean,average_log2Fold)) %>% 
  mutate(variable = ifelse(grepl("stmn_2",name), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = tissue_clean, y = value,fill = name)) +
  geom_boxplot() +
  geom_point(pch = 21,height = 0,width = 0.2,position = position_jitterdodge()) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean+disease_group2,nrow = 1,scales = 'free')  +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')
```

## Side by side comparison UNC13A & STMN2 CE PSI - v3
```{r}
clean_data_table %>% 
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  # filter(unc13a_cryptic_leaf_psi > 0) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  melt() %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>% 
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable)) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  scale_color_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean+disease_group2,nrow = 3,scales = 'free') +
    scale_y_continuous(labels = scales::percent)  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')

```

## Log2FC UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>% 
  filter(UNC13A_annotated_leaf > 10) %>% 
  filter(STMN2_annotated_leaf > 10) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 

  # mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>% 
  ggplot(aes(x = tissue_clean,y = log2_unc, fill = tissue_clean)) + 
  geom_boxplot(position = 'dodge') +
  # geom_jitter(height = 0,width = 0.2,pch = 21) +
  facet_wrap(~disease_group2,scales = "free_x") + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```





# Effect of UNC13A genotype

## UNC13A psi by genotype
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## UNC13A psi by genotype - tissue split
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## UNC13A psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```


## STMN2 psi by genotype 
```{r}

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## Correlation UNC13A PSI and STMN2 PSI with genotype effect

```{r}


clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>%
  filter(!grepl("Cord",tissue_clean)) %>% 
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'rs12973192',                      
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = rs12973192),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32))


```


## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex only

```{r}
detected_table_for_graph = clean_data_table %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))



comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F) 

detected_table_for_graph %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means( label = "p.format") + 
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) 
```



## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord =

```{r}
detected_table_for_graph = clean_data_table %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))



comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F) 

detected_table_for_graph %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means( label = "p.format") + 
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) 
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932

```{r}
detected_table_for_graph_rs12608932 = clean_data_table %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(rs12608932) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))



comparisons = combn(unique(detected_table_for_graph_rs12608932$detection_name),2,simplify = F) 

detected_table_for_graph_rs12608932 %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means( label = "p.format") + 
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12608932 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) 
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932

```{r}
clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(rs12608932) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means( label = "p.format") + 
  stat_compare_means(aes(group = detection_name),comparisons = detected_comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12608932 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) 
```

## Detection rate by genotype


```{r}
overall_fisher = clean_data_table %>%
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  dplyr::select(participant_id,unc13a_detected,rs12973192) %>% 
  unique() %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(unc13a_detected)) %>% 
  dplyr::select(rs12973192,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(n_non_detected = n_sample - n_sample_detected) %>% 
  dplyr::select(-n_sample)


over_p = overall_fisher %>% 
  column_to_rownames('rs12973192') %>% 
  chisq.test() %>% 
  broom::tidy() %>% 
  .$p.value

overall_fisher %>% 
  mutate(n_sample = n_non_detected + n_sample_detected) %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
  geom_col(show.legend = F) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(labels = scales::percent) + 
  ylab("% of TDP-43 Proteionopathy Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals") + 
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) 

```



Although this difference is not significant, with the Fisher's exact giving
a p-value of `r round(over_p,2)`.


# Variant allele expression in TDP-43 KD experiments

Across our KD's, we observed a clear allelic bias in the SK-DZ-N cells

```{r}
kd_vafs = fread(file.path(here::here(),"data","all_kds_unc13.snp.out.tsv"))

kd_vafs %>% 
    left_join(rel_rna_cryptic_amount) %>% 
    filter(!is.na(condition)) %>% 
    mutate(VAF = alt_reads / total) %>% 
    filter(condition != "control" ) %>% 
    filter(alt_reads > 5) %>%  
    ggplot(aes(x = stmn_2_cryptic_psi,y = cryptic_psi, fill = -(VAF - 0.5),size = total)) + 
    geom_point(pch = 21)  + 
    coord_fixed() +
  ylim(0,0.8) + 
  xlim(0,0.8) +
  geom_abline() + 
    scale_fill_gradient2()

```

# Liu Facs Neurons

Also look at the coverage in the Liu nuclear facs sorted neurons

```{r}
liu_vafs = fread(file.path(here::here(),"data","liu_facs_neurons_unc13.snp.out.tsv"))
liu_stmn2 = fread(file.path(here::here(),"data","liu_stmn2_and_unc13aaggregated.clean.annotated.bed"))
liu_stmn2[,sample := gsub(".SJ.out","",V4)]
liu_stmn2 = liu_stmn2 %>% 
  unique() %>% 
    pivot_wider(id_cols = 'sample',
                names_from = "V7",
                values_from = 'V5',
                values_fill = 0)

liu_stmn2 <- liu_stmn2 %>% 
   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>% 
   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2)) 
liu_vafs %>% 
    mutate(VAF = alt_reads / total) %>% 
    data.table::melt(id.vars = c("sample",'VAF','total')) %>% 
    mutate(allele = ifelse(variable == "ref_reads", "C","G")) %>% 
    mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
    mutate(sample_name = gsub("_unsorted", "", sample_name)) %>% 
    mutate(condition = sub(".*_", "", sample)) %>% 
    ggplot(aes(x = condition,y = value, fill = allele)) + 
    geom_col(pch = 21, size = 4) +
    facet_wrap(~sample_name,scales = "free") 
    

liu_stmn2 %>% 
  dplyr::select(stmn2_psi,sample,unc13a_psi) %>% 
  data.table::melt() %>% 
   mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
    filter(!grepl("unsorted",sample_name)) %>% 
    mutate(condition = sub(".*_", "", sample)) %>% 
    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>% 
    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>% 
    ggplot(aes(x = condition,y = value, fill = variable)) + 
    geom_col(size = 4,position = 'dodge') +
    facet_wrap(~sample_name,scales = "free_x") + 
    scale_fill_manual(values = c("#004D40","#882255")) +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    theme(text = element_text(size = 24))  +
      xlab("") +
    ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))

liu_stmn2 %>% 
   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>% 
   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime)) %>% 
   left_join(liu_vafs) %>% 
    mutate(alt_vaf = (alt_reads)/total) %>% 
    mutate(condition = ifelse(grepl("negative",sample),"negative", "positive")) %>% 
             filter(!grepl("unsorted",sample)) %>% 
             
    ggplot(aes(y = unc13a_psi,x = stmn2_psi, fill = alt_vaf)) + 
    geom_point(pch = 21,size = 5) + 
    ggtitle("STMN2 PSI and the VAF of the G allele \n in the Liu Nuclei ") +
  facet_grid(~condition) + 
  scale_fill_viridis_c(option = "plasma") + 
  coord_fixed() + 
  geom_abline()
```

## Variant allele expression in patients

```{r}
nycg_vafs = fread(file.path(here::here(),"data","NYGC_all_samples_UNC13A_snp_coverage.tsv"),fill = T)
nycg_vafs %>% 
    left_join(clean_data_table) %>% 
    filter(rs12973192 == "C/G") %>% 
    mutate(fraction_risk = alt_reads / total) %>% 
    filter(!is.na(fraction_risk)) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) + 
    xlab("STMN2 Cryptic PSI") +
    ylab("UNC13A CE PSI") +
    geom_abline() + 
    scale_fill_gradient2()

nycg_vafs %>% 
    left_join(clean_data_table) %>% 
    filter(rs12973192 == "C/G") %>% 
    mutate(fraction_risk = alt_reads / total) %>% 
    filter(!is.na(fraction_risk)) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    mutate(stmn2_unc_fract = ((unc13a_cryptic_leaf_psi * 100) / (stmn_2_cryptic_psi_leaf * 100))) %>% 
    ggplot(aes(x = stmn2_unc_fract, y = fraction_risk)) +
    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) + 
    xlab("STMN2 Cryptic PSI") +
    ylab("UNC13A CE PSI") +
    scale_fill_gradient2()
```


